MCC Model Predictions

### 2018-06-23 stats/model_predictions_variable_priors.Rmd
updating model comparison and predictions

@@

```{r bda model rsa noise, include = FALSE}
predictions.rsa.noise <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

var model = function(){
   var noise = 0.28

@@

  var conditionSpecificData =  _.filter(allData, {Experiment: conditionInfo.Experiment, Change: conditionInfo.Change, Alignment: conditionInfo.Alignment})

   var conditionSpecificPriors =  _.filter(allPriorProbs, {Experiment: conditionInfo.Experiment, Change: conditionInfo.Change, Alignment: conditionInfo.Alignment})

  //display(JSON.stringify(conditionSpecificData[0].correct_inf))

@@

  return  [conditionInfo.Experiment + "_" + conditionInfo.Change + "_" +  conditionInfo.Alignment, Math.exp(noisyModelPredictions.score(1))]

@@

 return extend(_.fromPairs(conditionOutput))
}
'
```

@@

strongPriorProbs <- data.frame(
  Experiment = c("Preference","Preference","Preference","Preference","Novelty","Novelty","Novelty","Novelty"),
  Change = c("Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker"),
  Alignment = c("Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent"))

@@

mediumPriorProbs <- data.frame(
  Experiment = c("Preference","Preference","Preference","Preference","Novelty","Novelty","Novelty","Novelty"),
  Change = c("Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker"),
  Alignment = c("Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent"))

@@

weakPriorProbs <- data.frame(
  Experiment = c("Preference","Preference","Preference","Preference","Novelty","Novelty","Novelty","Novelty"),
  Change = c("Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker"),
  Alignment = c("Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent"))

weakPriorProbs$priorProbs = list(
    c(0.1428571, 0.1428571, 0.8571429),
    c(0.8571429, 0.8571429, 0.1428571),
    c(0.4107143, 0.4107143, 0.5892857),
    c(0.5892857, 0.5892857, 0.4107143),
    c(0.4166667, 0.4166667, 0.5833333), 
    c(0.5833333, 0.5833333, 0.4166667),
    c(0.475, 0.475, 0.5250000),
    c(0.5250000, 0.5250000, 0.475))
```

```{r predictions rsa noise strong prior, include = FALSE}
bda <- webppl(
  program_code = paste(rsaUtils, rsaModel2, predictions.rsa.noise, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = strongPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
)

@@

noisy.rsa.model.strong <- bda%>%
  separate(Parameter, into = c("Experiment", "Change", "Alignment"), sep="_")%>%
  mutate(Model="Noisy RSA Model - Strong Prior",
         Prior = "Strong",
         Adjustment = "Noise",
         Type = "RSA")%>%
  group_by(Model,Adjustment,Type,Experiment,Change,Alignment) %>%
  summarise(mean = estimate_mode(value), ci_lower = hdi_lower(value), ci_upper = hdi_upper(value))

```{r predictions rsa noise medium prior, include = FALSE}
bda <- webppl(
  program_code = paste(rsaUtils, rsaModel2, predictions.rsa.noise, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = mediumPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
)

noisy.rsa.model.medium <- bda%>%
  separate(Parameter, into = c("Experiment", "Change", "Alignment"), sep="_")%>%
  mutate(Model="Noisy RSA Model - Medium Prior",
         Prior = "Medium",
         Adjustment = "Noise",
         Type = "RSA")%>%
  group_by(Model,Adjustment,Type,Experiment,Change,Alignment) %>%
  summarise(mean = estimate_mode(value), ci_lower = hdi_lower(value), ci_upper = hdi_upper(value))

```{r predictions rsa noise weak prior, include = FALSE}
bda <- webppl(
  program_code = paste(rsaUtils, rsaModel2, predictions.rsa.noise, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = weakPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
)

noisy.rsa.model.weak <- bda%>%
  separate(Parameter, into = c("Experiment", "Change", "Alignment"), sep="_")%>%
  mutate(Model="Noisy RSA Model - Weak Prior",
         Prior = "Weak",
         Adjustment = "Noise",
         Type = "RSA")%>%
  group_by(Model,Adjustment,Type,Experiment,Change,Alignment) %>%
  summarise(mean = estimate_mode(value), ci_lower = hdi_lower(value), ci_upper = hdi_upper(value))

@@

pred <- bind_rows(
  noisy.rsa.model.strong,
  noisy.rsa.model.medium,
  noisy.rsa.model.weak
) %>% 
  ungroup() %>%
  mutate(Model = relevel(as.factor(Model), ref = "Noisy RSA Model - Weak Prior"))

ggplot(pred, 
       aes(x = Change, y = mean, fill = Alignment)) +
  geom_bar(stat="identity", position = position_dodge(), color = 'black') + 
  geom_hline(yintercept = 0.5, lty=2)+
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper),position = position_dodge(width = 0.9))+
  labs(x=" ",y="Proportion Choosing More Informative")+
  facet_grid(Experiment ~ Model)+
  theme_few(base_size = 12) + 
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  ylim(0,1)
```

>>>>

load("model.predictions.RData")

@@

```{r predictions rsa noise, include = FALSE}
predictions.rsa.noise <-'
var allConditions = dataFromR.newConditions
var allData = dataFromR.predData
var allPriorProbs = dataFromR.predPriorProbs

var model = function(){
   var noise = uniformDrift({a: 0, b:1, width: 0.1})

@@

  var conditionSpecificData =  _.filter(allData, {Experiment: conditionInfo.Experiment, Change: conditionInfo.Change, Alignment: conditionInfo.Alignment, Prior :"Strong"})

   var conditionSpecificPriors =  _.filter(allPriorProbs, {Experiment: conditionInfo.Experiment, Change: conditionInfo.Change, Alignment: conditionInfo.Alignment, Prior :"Strong"})

@@

  return  [conditionInfo.Experiment + "_" + conditionInfo.Change + "_" +  conditionInfo.Alignment+ "_" + conditionInfo.Prior, Math.exp(noisyModelPredictions.score(1))]

@@

 // return extend(_.fromPairs(conditionOutput), {noise: noise})

// generate predictions for new conditions

   var newPredictions = map(function(conditionInfo){

    var newSpecificPriors =  _.filter(allPriorProbs, {Experiment: conditionInfo.Experiment, Change: conditionInfo.Change, Alignment: conditionInfo.Alignment, Prior: conditionInfo.Prior})

    var newModelPredictions = pragmaticListener({label: "dax", point: 2 }, newSpecificPriors[0].priorProbs)
    var newNoisyModelPredictions = addNoise(newModelPredictions, noise)

    return  [conditionInfo.Experiment + "_" + conditionInfo.Change + "_" +  conditionInfo.Alignment + "_" + conditionInfo.Prior, Math.exp(newNoisyModelPredictions.score(1))]

  return extend(_.fromPairs(newPredictions), {noise: noise})

}
'
```

@@

strongPriorProbs <- data.frame(
  Experiment = c("Preference","Preference","Preference","Preference","Novelty","Novelty","Novelty","Novelty"),
  Change = c("Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker"),
  Alignment = c("Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent"),
  Prior = rep("Strong",8))

@@

mediumPriorProbs <- data.frame(
  Experiment = c("Preference","Preference","Preference","Preference","Novelty","Novelty","Novelty","Novelty"),
  Change = c("Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker"),
  Alignment = c("Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent"),
  Prior = rep("Medium",8))

@@

weakPriorProbs <- data.frame(
  Experiment = c("Novelty","Novelty","Novelty","Novelty"),
  Change = c("Same speaker","Same speaker","Different speaker","Different speaker"),
  Alignment = c("Congruent","Incongruent","Congruent","Incongruent"),
  Prior = rep("Weak",4))

weakPriorProbs$priorProbs = list(
  c(0.4166667, 0.4166667, 0.5833333), 
  c(0.5833333, 0.5833333, 0.4166667),
  c(0.475, 0.475, 0.5250000),
  c(0.5250000, 0.5250000, 0.475))

predPriorProbs = rbind(strongPriorProbs,mediumPriorProbs,weakPriorProbs)

allPriorProbs <- strongPriorProbs %>%
  select(-Prior)

allConditions <- predPriorProbs %>%
  filter(Prior == "Strong")%>%
  select(-priorProbs, -Prior)

newConditions <- predPriorProbs %>%
  select(-priorProbs)

predData <- data %>%
  mutate(Prior = "Strong")
```

```{r predictions rsa noise, include = FALSE}
bda.pred.noise <- webppl(
  program_code = paste(rsaUtils, rsaModel2, predictions.rsa.noise, sep='\n'),
  data = list(allData = data, newConditions = newConditions, predPriorProbs = predPriorProbs, predData = predData), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 10000, burn = 2000, verbose = T)
)

## distribution of noise
posterior.noise.pred <- bda.pred.noise %>%
  filter(Parameter %in% c("noise", "priorExponent"))

ggplot(posterior.noise.pred, aes(x = value))+
  geom_histogram()+
  facet_wrap(~Parameter, scales = 'free')

@@

noisy.rsa.pred <- bda.pred.noise %>%
  filter(!(Parameter %in% c("noise", "priorExponent")))  %>%
  separate(Parameter, into = c("Experiment", "Change", "Alignment", "Prior"), sep="_")%>%
  mutate(Model="Noisy RSA Model",
         Adjustment = "Noise",
         Type = "RSA",
         Prior = relevel(as.factor(Prior), ref = "Weak"))%>%
  group_by(Model,Prior,Adjustment,Type,Experiment,Change,Alignment) %>%
  summarise(mean = estimate_mode(value), ci_lower = hdi_lower(value), ci_upper = hdi_upper(value))

@@

ggplot(noisy.rsa.pred, 
       aes(x = Change, y = mean, fill = Alignment)) +
  geom_bar(stat="identity", position = position_dodge(), color = 'black') + 
  geom_hline(yintercept = 0.5, lty=2)+
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper),position = position_dodge(width = 0.9))+
  labs(x=" ",y="Proportion Choosing More Informative")+
  facet_grid(Experiment ~ Prior)+
  theme_few(base_size = 12) + 
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  ylim(0,1)
```

```{r}
save.image(file="model.predictions.RData")
```
